Automated residual plot assessment with the \textsf{R} package \textsf{autovi} and the \textsf{Shiny} application \textsf{autovi.web}

### 2025-09-07 autovi_paper.qmd
Use AU spellling

---
title: Automated residual plot assessment with the \textsf{R} package \textsf{autovi} and the \textsf{Shiny} application \textsf{autovi.web}
format:
  quarto-anzjs-pdf:
    keep-tex: true  
    shorttitle: "Automated Residual Plot Assessment"
    double-space: true
    line-numbers: true
    # You can specify extra preamble in the file below.
    include-in-header:
      - header.tex
date: today
date-format: "YYYY" # used to define volumeyear
author:
  - name: Weihao Li
    email: patrick.li@anu.edu.au
    corresponding: true
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
      - name: The Australian National University
        department: Biological Data Science Institute
        address: 46 Sullivan's Creek Road, ACT 2600, Australia
    orcid: 0000-0003-4959-106X
  - name: Dianne Cook
    email: dicook@monash.edu 
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0002-3813-7155
  - name: Emi Tanaka
    email: emi.tanaka@anu.edu.au
    affiliations: 
      - name: The Australian National University
        department: Biological Data Science Institute
        address: 46 Sullivan's Creek Road, ACT 2600, Australia
    orcid: 0000-0002-1455-259X
  - name: Susan VanderPlas
    email: emi.tanaka@anu.edu.au
    affiliations:
      - name: University of Nebraska
        department: Department of Statistics
        address: Hardin Hall, 3310 Holdrege St Suite 340, Lincoln, NE 68583, United States
    orcid: 0000-0002-3803-0972
  - name: Klaus Ackermann
    email: Klaus.Ackermann@monash.edu 
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0001-7693-8538
# Note that the Journal requires that a paper must begin with a
# "Summary" not an "Abstract".  This is automatically taken care
# of by the anzsauth document style.  So even though the following
# says "abstract" the heading "Summary" will appear in
# the processed version.
abstract: |
  Visual assessment of residual plots is a common approach for diagnosing linear models, but it relies on manual evaluation, which does not scale well and can lead to inconsistent decisions across analysts. The lineup protocol, which embeds the observed plot among null plots, can reduce subjectivity but requires even more human effort. In today’s data-driven world, such tasks are well-suited for automation. We present a new \textsf{R} package that uses a computer vision model to automate the evaluation of residual plots. An accompanying \textsf{Shiny} application is provided for ease of use. Given a sample of residuals, the model predicts a visual signal strength (VSS) and offers supporting information to help analysts assess model fit. 
# Note that "keywords" should not include words and phrases
# that form part of the title of the paper.
# Also keywords (with the exception of proper names and
# certain abbreviations that conventionally appear all in
# capital letters) should *not* be capitalised.
keywords: [initial data analysis, statistical graphics, data visualization, visual inference, computer vision, machine learning, hypothesis testing, regression analysis, model diagnostics]
bibliography: bibliography.bib  
#nocite: |
---

@@

Regression analysis is a widely used statistical modeling technique for data in many fields. There is a vast array of software for conducting regression modeling and generating diagnostics. The package \textsf{lmtest} [@lmtest] provides a suite of conventional tests, while the \textsf{stats} package [@stats] offers standard diagnostic plots such as residuals vs. fitted values, quantile-quantile (Q-Q) plots, and residuals vs. leverage plots. Additional packages like \textsf{jtools} [@jtools], \textsf{olsrr} [@olsrr], \textsf{rockchalk} [@rockchalk], and \textsf{ggResidpanel} [@ggresidpanel] deliver similar graphical diagnostics, often with enhanced aesthetics or interactive features. These tools collectively produce the core diagnostic plots outlined in the classical text by @cook1982residuals. The \textsf{ecostats} package [@warton_global_2023] extends these diagnostics by incorporating simulation envelopes into residual plots. Meanwhile, \textsf{DHARMa} [@dharma] compares empirical quantiles (0.25, 0.5, and 0.75) of scaled residuals to their theoretical counterparts, with a strong focus on identifying model violations such as heteroscedasticity, misspecified functional forms, and issues specific to generalized linear and mixed-effect models, like over/under-dispersion. It also provides conventional test annotations to reduce the risk of misinterpretation.

@@

In the context of classical normal linear regression model diagnostics, VSS describes the clarity of visual patterns on a diagnostic plot that may indicate model violations. Violations can be categorized as weak, moderate, or strong, but here we treat it as a continuous positive real variable. Importantly, its interpretation depends on how it is linked to a function of the data or the underlying data generating process. Consequently, the calculation of VSS can vary across different model classes or within the same model, depending on the generating function.

@@

The computer vision model approximates this mapping from a set of residuals to its corresponding distance measure. It is trained on a large number of synthetic regression models, each designed to simulate specific violations of classical linear regression assumptions. These models incorporate non-linearity through Hermite polynomial transformations of predictors, heteroskedasticity by making the error variance a predictor-dependent function, and non-normality by drawing residuals from distributions such as discrete, uniform, and lognormal. Both simple and multiple linear regression structures are used, with controlled parameters to generate diverse and complex residual patterns. Since the data-generating process is known, the distance measure $D$ can be explicitly calculated, enabling supervised training. The computer vision model takes a residual plot as input and outputs the corresponding distance measure, learning to quantify model violations directly from visual patterns. Additional details are provided in @li2024automated.

@@

```{r}
#| label: fig-three-examples
#| fig-cap: "Reading residual plots can be a difficult task, particularly for students new to statistical modeling. The \\textsf{autovi} package makes it easier. Here are three examples of residual plots, which may appear to have structure. According to autovi, the visual signal strengths (VSS) of these three examples are approximately (a) 1.53, (b) 3.57, (c) 5.87, resulting in (b), (c) being significant violations of good residuals, but (a) is consistent with a good residual plot."
#| fig-width: 12
#| fig-height: 4
#| out-width: 100%
library(readr)
library(dplyr)
library(ggplot2)
library(patchwork)
library(ggthemes)
library(bandicoot)

@@

The package offers a straightforward visualization of the assessment result through the `summary_plot()` function.

@@

The initial motivation for developing \textsf{autovi} was to create a convenient interface for sharing the models described and trained in @li2024automated. However, recognizing that the classical normal linear regression model represents a restricted class of models, we sought to avoid limiting the potential for future extensions, whether by the original developers or other developers. As a result, the package was designed to function seamlessly with linear regression models with minimal modification and few required arguments, while also accommodating other classes of models through partial infrastructure substitution. This modular and customisable design allows \textsf{autovi} to handle a wide range of residual diagnostics tasks.

The infrastructure of \textsf{autovi} consists of ten core modules: data extraction, bootstrapping and model refitting, fitted values and residuals extraction, auxiliary computation, null residual simulation, plotting, plot saving, image reading and resizing, VSS prediction, and $p$-value computation. Each module is designed with minimal dependency on the preceding modules, allowing users to customize parts of the infrastructure without affecting its overall integrity. An overview of this infrastructure is illustrated in @fig-autovi-diag.

@@

The \textsf{autovi} \textsf{R} package can be extended to accommodate other classes of models beyond linear regression, such as generalized linear models (`glm`). This is achieved by substituting the relevant overridable modules, and if needed, supplying a different \textsf{Keras} model.

@@

```{r fig-pois-lineup, fig.cap="A lineup of residual plots from Poisson generalized linear models, with the true residual plot at position 4, which displays a distinct U-shaped pattern. In contrast, the null plots show characteristics broadly consistent with well-behaved residuals from linear regression models.", fig.pos="H", out.width="80%"}
pois_checker$plot_lineup()
```

@@

While \textsf{autovi} requires downloading the pre-trained computer vision models from GitHub, these models in '.keras' file format are incompatible with \textsf{TensorFlow.js}. Therefore, we extract and store the model weights in JSON files and include them as extra resources in the \textsf{Shiny} application. When the application initializes, \textsf{TensorFlow.js} rebuilds the computer vision model using these pre-stored weights.

To allow communication between \textsf{TensorFlow.js} and other components of the \textsf{Shiny} application, the \textsf{shinyjs} \textsf{R} package [@shinyjs] is used. This package allows calling custom \textsf{JavaScript} code within the \textsf{Shiny} framework. The specialized \textsf{JavaScript} code for initializing \textsf{TensorFlow.js} and calling \textsf{TensorFlow.js} for VSS prediction is deployed alongside the \textsf{Shiny} application as additional resources.

@@

The framework established by \textsf{autovi} and \textsf{autovi.web} opens up exciting possibilities for further research and development. Future work could explore the extension of these automated assessment techniques to other types of diagnostic plots and statistical models, potentially revolutionizing how we approach statistical inference using visual displays more broadly.

>>>>

---
title: Automated residual plot assessment with the \textsf{R} package \textsf{autovi} and the \textsf{Shiny} application \textsf{autovi.web}
format:
  quarto-anzjs-pdf:
    keep-md: true
    keep-tex: true  
    shorttitle: "Automated Residual Plot Assessment"
    double-space: true
    line-numbers: true
    # You can specify extra preamble in the file below.
    include-in-header:
      - header.tex
date: today
date-format: "YYYY" # used to define volumeyear
author:
  - name: Weihao Li
    email: patrick.li@anu.edu.au
    corresponding: true
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
      - name: The Australian National University
        department: Biological Data Science Institute
        address: 46 Sullivan's Creek Road, ACT 2600, Australia
    orcid: 0000-0003-4959-106X
  - name: Dianne Cook
    email: dicook@monash.edu 
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0002-3813-7155
  - name: Emi Tanaka
    email: emi.tanaka@anu.edu.au
    affiliations: 
      - name: The Australian National University
        department: Biological Data Science Institute
        address: 46 Sullivan's Creek Road, ACT 2600, Australia
    orcid: 0000-0002-1455-259X
  - name: Susan VanderPlas
    email: emi.tanaka@anu.edu.au
    affiliations:
      - name: University of Nebraska
        department: Department of Statistics
        address: Hardin Hall, 3310 Holdrege St Suite 340, Lincoln, NE 68583, United States
    orcid: 0000-0002-3803-0972
  - name: Klaus Ackermann
    email: Klaus.Ackermann@monash.edu 
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0001-7693-8538
# Note that the Journal requires that a paper must begin with a
# "Summary" not an "Abstract".  This is automatically taken care
# of by the anzsauth document style.  So even though the following
# says "abstract" the heading "Summary" will appear in
# the processed version.
abstract: |
  Visual assessment of residual plots is a common approach for diagnosing linear models, but it relies on manual evaluation, which does not scale well and can lead to inconsistent decisions across analysts. The lineup protocol, which embeds the observed plot among null plots, can reduce subjectivity but requires even more human effort. In today’s data-driven world, such tasks are well-suited for automation. We present a new \textsf{R} package that uses a computer vision model to automate the evaluation of residual plots. An accompanying \textsf{Shiny} application is provided for ease of use. Given a sample of residuals, the model predicts a visual signal strength (VSS) and offers supporting information to help analysts assess model fit. 
# Note that "keywords" should not include words and phrases
# that form part of the title of the paper.
# Also keywords (with the exception of proper names and
# certain abbreviations that conventionally appear all in
# capital letters) should *not* be capitalised.
keywords: [initial data analysis, statistical graphics, data visualisation, visual inference, computer vision, machine learning, hypothesis testing, regression analysis, model diagnostics]
bibliography: bibliography.bib  
#nocite: |
---

@@

Regression analysis is a widely used statistical modelling technique for data in many fields. There is a vast array of software for conducting regression modelling and generating diagnostics. The package \textsf{lmtest} [@lmtest] provides a suite of conventional tests, while the \textsf{stats} package [@stats] offers standard diagnostic plots such as residuals vs. fitted values, quantile-quantile (Q-Q) plots, and residuals vs. leverage plots. Additional packages like \textsf{jtools} [@jtools], \textsf{olsrr} [@olsrr], \textsf{rockchalk} [@rockchalk], and \textsf{ggResidpanel} [@ggresidpanel] deliver similar graphical diagnostics, often with enhanced aesthetics or interactive features. These tools collectively produce the core diagnostic plots outlined in the classical text by @cook1982residuals. The \textsf{ecostats} package [@warton_global_2023] extends these diagnostics by incorporating simulation envelopes into residual plots. Meanwhile, \textsf{DHARMa} [@dharma] compares empirical quantiles (0.25, 0.5, and 0.75) of scaled residuals to their theoretical counterparts, with a strong focus on identifying model violations such as heteroscedasticity, misspecified functional forms, and issues specific to generalised linear and mixed-effect models, like over/under-dispersion. It also provides conventional test annotations to reduce the risk of misinterpretation.

@@

In the context of classical normal linear regression model diagnostics, VSS describes the clarity of visual patterns on a diagnostic plot that may indicate model violations. Violations can be categorised as weak, moderate, or strong, but here we treat it as a continuous positive real variable. Importantly, its interpretation depends on how it is linked to a function of the data or the underlying data generating process. Consequently, the calculation of VSS can vary across different model classes or within the same model, depending on the generating function.

@@

The computer vision model approximates this mapping from a set of residuals to its corresponding distance measure. It is trained on a large number of synthetic regression models, each designed to simulate specific violations of classical linear regression assumptions. These models incorporate non-linearity through Hermite polynomial transformations of predictors, heteroscedasticity by making the error variance a predictor-dependent function, and non-normality by drawing residuals from distributions such as discrete, uniform, and lognormal. Both simple and multiple linear regression structures are used, with controlled parameters to generate diverse and complex residual patterns. Since the data-generating process is known, the distance measure $D$ can be explicitly calculated, enabling supervised training. The computer vision model takes a residual plot as input and outputs the corresponding distance measure, learning to quantify model violations directly from visual patterns. Additional details are provided in @li2024automated.

@@

```{r}
#| label: fig-three-examples
#| fig-cap: "Reading residual plots can be a difficult task, particularly for students new to statistical modelling. The \\textsf{autovi} package makes it easier. Here are three examples of residual plots, which may appear to have structure. According to autovi, the visual signal strengths (VSS) of these three examples are approximately (a) 1.53, (b) 3.57, (c) 5.87, resulting in (b), (c) being significant violations of good residuals, but (a) is consistent with a good residual plot."
#| fig-width: 12
#| fig-height: 4
#| out-width: 100%
library(readr)
library(dplyr)
library(ggplot2)
library(patchwork)
library(ggthemes)
library(bandicoot)

@@

The package offers a straightforward visualisation of the assessment result through the `summary_plot()` function.

@@

The initial motivation for developing \textsf{autovi} was to create a convenient interface for sharing the models described and trained in @li2024automated. However, recognising that the classical normal linear regression model represents a restricted class of models, we sought to avoid limiting the potential for future extensions, whether by the original developers or other developers. As a result, the package was designed to function seamlessly with linear regression models with minimal modification and few required arguments, while also accommodating other classes of models through partial infrastructure substitution. This modular and customisable design allows \textsf{autovi} to handle a wide range of residual diagnostics tasks.

The infrastructure of \textsf{autovi} consists of ten core modules: data extraction, bootstrapping and model refitting, fitted values and residuals extraction, auxiliary computation, null residual simulation, plotting, plot saving, image reading and resizing, VSS prediction, and $p$-value computation. Each module is designed with minimal dependency on the preceding modules, allowing users to customise parts of the infrastructure without affecting its overall integrity. An overview of this infrastructure is illustrated in @fig-autovi-diag.

@@

The \textsf{autovi} \textsf{R} package can be extended to accommodate other classes of models beyond linear regression, such as generalised linear models (`glm`). This is achieved by substituting the relevant overridable modules, and if needed, supplying a different \textsf{Keras} model.

@@

```{r fig-pois-lineup, fig.cap="A lineup of residual plots from Poisson generalised linear models, with the true residual plot at position 4, which displays a distinct U-shaped pattern. In contrast, the null plots show characteristics broadly consistent with well-behaved residuals from linear regression models.", fig.pos="H", out.width="80%"}
pois_checker$plot_lineup()
```

@@

While \textsf{autovi} requires downloading the pre-trained computer vision models from GitHub, these models in '.keras' file format are incompatible with \textsf{TensorFlow.js}. Therefore, we extract and store the model weights in JSON files and include them as extra resources in the \textsf{Shiny} application. When the application initialises, \textsf{TensorFlow.js} rebuilds the computer vision model using these pre-stored weights.

To allow communication between \textsf{TensorFlow.js} and other components of the \textsf{Shiny} application, the \textsf{shinyjs} \textsf{R} package [@shinyjs] is used. This package allows calling custom \textsf{JavaScript} code within the \textsf{Shiny} framework. The specialised \textsf{JavaScript} code for initialising \textsf{TensorFlow.js} and calling \textsf{TensorFlow.js} for VSS prediction is deployed alongside the \textsf{Shiny} application as additional resources.

@@

The framework established by \textsf{autovi} and \textsf{autovi.web} opens up exciting possibilities for further research and development. Future work could explore the extension of these automated assessment techniques to other types of diagnostic plots and statistical models, potentially revolutionising how we approach statistical inference using visual displays more broadly.

### 2025-09-07 autovi_paper.tex
Use AU spellling

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Automated residual plot assessment with the  package  and the  application },
  pdfauthor={Weihao Li; Dianne Cook; Emi Tanaka; Susan VanderPlas; Klaus Ackermann},
  pdfkeywords={initial data analysis, statistical graphics, data
visualization, visual inference, computer vision, machine
learning, hypothesis testing, regression analysis, model diagnostics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\keywords{initial data analysis; statistical graphics; data
visualization; visual inference; computer vision; machine
learning; hypothesis testing; regression analysis; model diagnostics}

@@

Regression analysis is a widely used statistical modeling technique for
data in many fields. There is a vast array of software for conducting
regression modeling and generating diagnostics. The package
\textsf{lmtest} \citep{lmtest} provides a suite of conventional tests,
while the \textsf{stats} package \citep{stats} offers standard
diagnostic plots such as residuals vs.~fitted values, quantile-quantile
(Q-Q) plots, and residuals vs.~leverage plots. Additional packages like
\textsf{jtools} \citep{jtools}, \textsf{olsrr} \citep{olsrr},
\textsf{rockchalk} \citep{rockchalk}, and \textsf{ggResidpanel}
\citep{ggresidpanel} deliver similar graphical diagnostics, often with
enhanced aesthetics or interactive features. These tools collectively
produce the core diagnostic plots outlined in the classical text by
\citet{cook1982residuals}. The \textsf{ecostats} package
\citep{warton_global_2023} extends these diagnostics by incorporating
simulation envelopes into residual plots. Meanwhile, \textsf{DHARMa}
\citep{dharma} compares empirical quantiles (0.25, 0.5, and 0.75) of
scaled residuals to their theoretical counterparts, with a strong focus
on identifying model violations such as heteroscedasticity, misspecified
functional forms, and issues specific to generalized linear and
mixed-effect models, like over/under-dispersion. It also provides
conventional test annotations to reduce the risk of misinterpretation.

@@

The computer vision model approximates this mapping from a set of
residuals to its corresponding distance measure. It is trained on a
large number of synthetic regression models, each designed to simulate
specific violations of classical linear regression assumptions. These
models incorporate non-linearity through Hermite polynomial
transformations of predictors, heteroskedasticity by making the error
variance a predictor-dependent function, and non-normality by drawing
residuals from distributions such as discrete, uniform, and lognormal.
Both simple and multiple linear regression structures are used, with
controlled parameters to generate diverse and complex residual patterns.
Since the data-generating process is known, the distance measure \(D\)
can be explicitly calculated, enabling supervised training. The computer
vision model takes a residual plot as input and outputs the
corresponding distance measure, learning to quantify model violations
directly from visual patterns. Additional details are provided in
\citet{li2024automated}.

@@

\caption{\label{fig-three-examples}Reading residual plots can be a
difficult task, particularly for students new to statistical modeling.
The \textsf{autovi} package makes it easier. Here are three examples of
residual plots, which may appear to have structure. According to autovi,
the visual signal strengths (VSS) of these three examples are
approximately (a) 1.53, (b) 3.57, (c) 5.87, resulting in (b), (c) being
significant violations of good residuals, but (a) is consistent with a
good residual plot.}

@@

The package offers a straightforward visualization of the assessment
result through the \texttt{summary\_plot()} function.

@@

The initial motivation for developing \textsf{autovi} was to create a
convenient interface for sharing the models described and trained in
\citet{li2024automated}. However, recognizing that the classical normal
linear regression model represents a restricted class of models, we
sought to avoid limiting the potential for future extensions, whether by
the original developers or other developers. As a result, the package
was designed to function seamlessly with linear regression models with
minimal modification and few required arguments, while also
accommodating other classes of models through partial infrastructure
substitution. This modular and customisable design allows
\textsf{autovi} to handle a wide range of residual diagnostics tasks.

The infrastructure of \textsf{autovi} consists of ten core modules: data
extraction, bootstrapping and model refitting, fitted values and
residuals extraction, auxiliary computation, null residual simulation,
plotting, plot saving, image reading and resizing, VSS prediction, and
\(p\)-value computation. Each module is designed with minimal dependency
on the preceding modules, allowing users to customize parts of the
infrastructure without affecting its overall integrity. An overview of
this infrastructure is illustrated in Figure~\ref{fig-autovi-diag}.

@@

The \textsf{autovi} \textsf{R} package can be extended to accommodate
other classes of models beyond linear regression, such as generalized
linear models (\texttt{glm}). This is achieved by substituting the
relevant overridable modules, and if needed, supplying a different
\textsf{Keras} model.

@@

\caption{\label{fig-pois-lineup}A lineup of residual plots from Poisson
generalized linear models, with the true residual plot at position 4,
which displays a distinct U-shaped pattern. In contrast, the null plots
show characteristics broadly consistent with well-behaved residuals from
linear regression models.}

@@

While \textsf{autovi} requires downloading the pre-trained computer
vision models from GitHub, these models in `.keras' file format are
incompatible with \textsf{TensorFlow.js}. Therefore, we extract and
store the model weights in JSON files and include them as extra
resources in the \textsf{Shiny} application. When the application
initializes, \textsf{TensorFlow.js} rebuilds the computer vision model
using these pre-stored weights.

To allow communication between \textsf{TensorFlow.js} and other
components of the \textsf{Shiny} application, the \textsf{shinyjs}
\textsf{R} package \citep{shinyjs} is used. This package allows calling
custom \textsf{JavaScript} code within the \textsf{Shiny} framework. The
specialized \textsf{JavaScript} code for initializing
\textsf{TensorFlow.js} and calling \textsf{TensorFlow.js} for VSS
prediction is deployed alongside the \textsf{Shiny} application as
additional resources.

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Automated residual plot assessment with the  package  and the  application },
  pdfauthor={Weihao Li; Dianne Cook; Emi Tanaka; Susan VanderPlas; Klaus Ackermann},
  pdfkeywords={initial data analysis, statistical graphics, data
visualisation, visual inference, computer vision, machine
learning, hypothesis testing, regression analysis, model diagnostics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\keywords{initial data analysis; statistical graphics; data
visualisation; visual inference; computer vision; machine
learning; hypothesis testing; regression analysis; model diagnostics}

@@

Regression analysis is a widely used statistical modelling technique for
data in many fields. There is a vast array of software for conducting
regression modelling and generating diagnostics. The package
\textsf{lmtest} \citep{lmtest} provides a suite of conventional tests,
while the \textsf{stats} package \citep{stats} offers standard
diagnostic plots such as residuals vs.~fitted values, quantile-quantile
(Q-Q) plots, and residuals vs.~leverage plots. Additional packages like
\textsf{jtools} \citep{jtools}, \textsf{olsrr} \citep{olsrr},
\textsf{rockchalk} \citep{rockchalk}, and \textsf{ggResidpanel}
\citep{ggresidpanel} deliver similar graphical diagnostics, often with
enhanced aesthetics or interactive features. These tools collectively
produce the core diagnostic plots outlined in the classical text by
\citet{cook1982residuals}. The \textsf{ecostats} package
\citep{warton_global_2023} extends these diagnostics by incorporating
simulation envelopes into residual plots. Meanwhile, \textsf{DHARMa}
\citep{dharma} compares empirical quantiles (0.25, 0.5, and 0.75) of
scaled residuals to their theoretical counterparts, with a strong focus
on identifying model violations such as heteroscedasticity, misspecified
functional forms, and issues specific to generalised linear and
mixed-effect models, like over/under-dispersion. It also provides
conventional test annotations to reduce the risk of misinterpretation.

@@

The computer vision model approximates this mapping from a set of
residuals to its corresponding distance measure. It is trained on a
large number of synthetic regression models, each designed to simulate
specific violations of classical linear regression assumptions. These
models incorporate non-linearity through Hermite polynomial
transformations of predictors, heteroscedasticity by making the error
variance a predictor-dependent function, and non-normality by drawing
residuals from distributions such as discrete, uniform, and lognormal.
Both simple and multiple linear regression structures are used, with
controlled parameters to generate diverse and complex residual patterns.
Since the data-generating process is known, the distance measure \(D\)
can be explicitly calculated, enabling supervised training. The computer
vision model takes a residual plot as input and outputs the
corresponding distance measure, learning to quantify model violations
directly from visual patterns. Additional details are provided in
\citet{li2024automated}.

@@

\caption{\label{fig-three-examples}Reading residual plots can be a
difficult task, particularly for students new to statistical modelling.
The \textsf{autovi} package makes it easier. Here are three examples of
residual plots, which may appear to have structure. According to autovi,
the visual signal strengths (VSS) of these three examples are
approximately (a) 1.53, (b) 3.57, (c) 5.87, resulting in (b), (c) being
significant violations of good residuals, but (a) is consistent with a
good residual plot.}

@@

The package offers a straightforward visualisation of the assessment
result through the \texttt{summary\_plot()} function.

@@

The initial motivation for developing \textsf{autovi} was to create a
convenient interface for sharing the models described and trained in
\citet{li2024automated}. However, recognising that the classical normal
linear regression model represents a restricted class of models, we
sought to avoid limiting the potential for future extensions, whether by
the original developers or other developers. As a result, the package
was designed to function seamlessly with linear regression models with
minimal modification and few required arguments, while also
accommodating other classes of models through partial infrastructure
substitution. This modular and customisable design allows
\textsf{autovi} to handle a wide range of residual diagnostics tasks.

The infrastructure of \textsf{autovi} consists of ten core modules: data
extraction, bootstrapping and model refitting, fitted values and
residuals extraction, auxiliary computation, null residual simulation,
plotting, plot saving, image reading and resizing, VSS prediction, and
\(p\)-value computation. Each module is designed with minimal dependency
on the preceding modules, allowing users to customise parts of the
infrastructure without affecting its overall integrity. An overview of
this infrastructure is illustrated in Figure~\ref{fig-autovi-diag}.

@@

The \textsf{autovi} \textsf{R} package can be extended to accommodate
other classes of models beyond linear regression, such as generalised
linear models (\texttt{glm}). This is achieved by substituting the
relevant overridable modules, and if needed, supplying a different
\textsf{Keras} model.

@@

\caption{\label{fig-pois-lineup}A lineup of residual plots from Poisson
generalised linear models, with the true residual plot at position 4,
which displays a distinct U-shaped pattern. In contrast, the null plots
show characteristics broadly consistent with well-behaved residuals from
linear regression models.}

@@

While \textsf{autovi} requires downloading the pre-trained computer
vision models from GitHub, these models in `.keras' file format are
incompatible with \textsf{TensorFlow.js}. Therefore, we extract and
store the model weights in JSON files and include them as extra
resources in the \textsf{Shiny} application. When the application
initialises, \textsf{TensorFlow.js} rebuilds the computer vision model
using these pre-stored weights.

To allow communication between \textsf{TensorFlow.js} and other
components of the \textsf{Shiny} application, the \textsf{shinyjs}
\textsf{R} package \citep{shinyjs} is used. This package allows calling
custom \textsf{JavaScript} code within the \textsf{Shiny} framework. The
specialised \textsf{JavaScript} code for initialising
\textsf{TensorFlow.js} and calling \textsf{TensorFlow.js} for VSS
prediction is deployed alongside the \textsf{Shiny} application as
additional resources.
